---
title: "Test cafe model and filter CAFE reusults for immune related"
author: "Kelly Ren"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2: default
  ---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


```{r}
#------------------------------------------------------
# Program name: Test_cafe_model_filter_for_immune.Rmd
# Objective: Differnet model avaliable when running cafe. The 2*(global_lnL−multi_lnL) was built to prove the correct lambda. The second part collect the CAFE results and filter for immune genes.
# Author: Lloyd Low and Kelly Ren
# Email add: lloydlow@hotmail.com
#------------------------------------------------------
```



```{r}
library(ggplot2)
library(easyGgplot2)
library(magrittr)
library(readr)
library(dplyr)
library(stringr)
library(tidyverse)
library(readxl)
library(gtools)
library(tibble)
library(reshape2)
library(GenomicFeatures)

'%!in%' <- function(x,y)!('%in%'(x,y))
```

# Test different model

Plot the distribution of 2*(global_lnL−multi_lnL) and prove the global lambda is the best to use.
CAFE provideed R cript lhtest.R to plot the distribution https://iu.app.box.com/v/cafetutorial-files/folder/22161185988.
Instead, can run following code to test the best model.

```{r}
lh2to1 <- read.table("Input/From_CAFE/lh2to1.out",header = F)%>%
  set_colnames(c("logLglobal", "lambda_global", "logLmultiple","lambda_multiple","lambda"))

lh2to1$diff <- 2*(lh2to1$logLglobal - lh2to1$logLmultiple)
#lh2to1 <- lh2to1[!is.infinite(lh2to1$diff),]
lh2to1 <- lh2to1[,c(1,3,6)]
global.lnk <- as.numeric(-44649.301048) # From 
multi.lnk <- as.numeric(-42342.085854)

# distribution of log likelyhood different/100
ggplot(lh2to1, aes(x=diff/100)) + 
  geom_histogram() + 
  xlim(-12, 12) +
  xlab("2*(global_lnL-multi_lnL)") + ylab("Count") +
    theme_bw()

2*(global.lnk-multi.lnk)/100
#Note that the observed likelihood ratio (2×(lnLglobal−lnLmulti)) would fall on the far left tail of the null distribution, yielding a very small p-value, and meaning that thethe probability of a multi-λmodel fitting better than a global-λmodel by chance is verysmall.
```

# error correction

```{r}
caferror_default_output <- read.table("Input/From_CAFE/caferror_default_output.txt",header = T)
min_Score <- min(caferror_default_output$Score)
subset(caferror_default_output, Score %in% min_Score) 
# The min score of caferror_default_output corresponding to the best modle to use for error correction
# So we go back to CAFE error contro result find the file cafe_errormodel_0.179832458496.txt as error model
```
